---
title: "Malurus Project"
author: "Elroy Au"
date: "12/02/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

### ================================== ###
### Data Analysis                      ###
### ================================== ###

### ================================== ###
### Principal Component Analysis       ###
### ================================== ###

```This is the code used to calculate the principal components used for future analyses and the code used to plot a PCA and a biplot from the data``` 

```{r} 

# load libraries 
library(ggplot2)
library(ggrepel)
library(scatterpie)   # for creating scatterpie graphs
library(readxl) 
library(RcppCNPy)     # for reading .npy (NUMPY) files generated in PCAngsd 
library(sf)           # for reading shapefiles 
library(readxl)

# load data
genetic.matrix <- npyLoad("genetic.matrix.cov.npy") # this file should contain the genetic covariance matrix of all of the unfiltered individuals, n=133
snp.filtered.genetic.matrix <- npyLoad("genetic.matrix.snp.filtered.cov.npy") # this file should contain the genetic covariance matrix of all the individuals post pca-filtering, n = 119
clim.filtered.genetic.matrix <- npyLoad("genetic.matrix.clim.filtered.cov.npy") # this file should contain the genetic covariance matrix of all of the individuals post-climate, n = 113
specimen.data <- read_excel("specimen-metadata.xlsx") # this file should contain the metadata for all of the sampled specimens

# analysis
# calculate eigenvalues
eigenvalues <- eigen(clim.filtered.genetic.matrix)$values         
# calculate eigenvectors
eigenvectors <- eigen(clim.filtered.genetic.matrix)$vectors  
# calculate the principal components using a matrix multiplication 
PC <- as.matrix(clim.filtered.genetic.matrix) %*% eigenvectors
# transpose the data in order to get the loading of the variables for each principal component 
PCs <- t(PC)

# take the principal component scores for individuals to create the PCA plot
PC1 <- PC[,1]
PC2 <- PC[,2]
PC3 <- PC[,3]

# transform into a dataframe
# note: principal components copied into metadata for further data analysis
PCA <- data.frame(PC1, PC2, PC3)

# plot PCA [1] collection state, collection date
ggplot(specimen.data, aes(x=PC1, y=PC2)) + geom_point(aes(shape = Period, color = STATE), position = "jitter", size = 2.5) + scale_shape_manual(values = c(3, 4, 8, 15, 17, 18, 19, 25)) + theme_bw()

# plot PCA [2] collection state, individual missingness
ggplot (specimen.data, aes(x=PC1, y=PC2)) + geom_point(aes(shape = Period, color = MISSINGNESS), position = "jitter", size = 2.5) + scale_shape_manual(values = c(3, 4, 8, 15, 17, 18, 19, 25)) + scale_color_continuous(high = "black", low = "turquoise2") + theme_bw() 

# plot biplot
# set pch and colour groups (grouped by collection state: NSW, NT, QLD, SA, VIC, WA, NA)
pch.group <- c(rep(21, times=10), rep(22, times=10), rep(23, times=10), rep(24, times=45), rep(3, times=1), rep(25, times=43))
# red = NSW, gold = NT, green = QLD, blue = SA, black = VIC, pink = WA, black X = N.A
col.group <- c(rep("red", times=10), rep("gold", times=10), rep("green", times=10), rep("skyblue2", times=45), rep("black", times=1), rep("violet", times=43))
# plot the individuals in principal component space 
plot(x=specimen.data$PC1, y=specimen.data$PC2, xlab="PC1 (3.18%)", ylab="PC2 (1.23%)", xlim=c(-1.0,1.0),ylim=c(-0.6,0.6), pch=pch.group, col="black", bg=col.group, cex=2, las=1, asp=1)
# y axis
abline(v=0, lty=2, col="grey50")
# x axis
abline(h=0, lty=2, col="grey50")
# set the x and y co-ordinates of the variables, which are equivalent to
# the loading scores of the variables (i.e. the individual samples)
# these will form the arrows on the bi-lot and were created by transposing the 
# principal component data made above (check PCs)
l.x <- PCs[,1]
l.y <- PCs[,2]
arrows(x0=0, x1=l.x, y0=0, y1=l.y, col="red", length = 0.1, lwd = 1.5)
# set text labels 
l.pos <- l.y
lo <- which(l.y < 0)
hi <- which(l.y > 0)
l.pos <- replace(l.pos, lo, "1")
l.pos <- replace(l.pos, hi, "3")
text(l.x, l.y, labels=specimen.data$`SAMPLE ID`, col="red", pos=l.pos)

# plot PCA on a geographic map
australia <- st_read("australia.shp")
# set objects
a <- ggplot() + geom_sf(data = australia, color = "black", fill = "white")
a + geom_point(data=specimen.data, aes(x=LONGITUDE, y=LATITUDE, color = PC1), size = 2.5) + scale_color_gradient2()
```

### ================================== ###
### ADMIXTURE ANALYSIS                 ###
### ================================== ###

```This is the code used to analyse the admixture proportions in the sample based off a hypothetical number of ancestral populations k```

```{r}

# load libraries
library(RColorBrewer)

# load new data
pop <- read.table("admix-pop-data.txt", fill = TRUE, header = FALSE) # this file contains the necessary metadata for ordering individuals via their collection state
q2 <- read.table("admix.2.txt", fill = TRUE, header = FALSE) # this file should contain the estimated admixture proportions for all individuals when the number of ancestral populations k = 2
q3 <- read.table("admix.3.txt", fill = TRUE, header = FALSE) # this file should contain the estimated admixture proportions for all individuals when the number of ancestral populations k = 3
q4 <- read.table("admix.4.txt", fill = TRUE, header = FALSE) # this file should contain the estimated admixture proportions for all individuals when the number of ancestral populations k = 4
q5 <- read.table("admix.5.txt", fill = TRUE, header = FALSE) # this file should contain the estimated admixture proportions for all individuals when the number of ancestral populations k = 5
q6 <- read.table("admix.6.txt", fill = TRUE, header = FALSE) # this file should contain the estimated admixture proportions for all individuals when the number of ancestral populations k = 6
# order according to population
ord <- order (pop[,2])
# plot admixture proportions 
barplot(t(q6)[,ord],col=brewer.pal(n=6, name="RdBu"), space=0, border=NA, xlab="Individuals", ylab="Admixture Proportions (K=6)")
# plot admixture proportions on a geographic map 





